---
title: "Happiness around the World (2015)"
author: "Josh Stremmel, Jooan Choi, Namrata Singh"
format:
  pdf:
    toc: true
csl: apa7.csl #or mla9.csl
execute:
  echo: false
---

## Overview

The World Happiness Report puts a number on something we all care about but rarely measure: how happy people are, country by country. Our project digs into the 2015 report to see what really drives those scores— is it money, health, freedom, or something else? By playing with the data we want to tell a short, visual story about what separates the happiest places from the not-so-happy ones, and maybe learn what our own country could do better. It’s basically a data-powered vibe check for the whole planet.

## Data Import
```{r}
library(tidyverse)
library(janitor)

happy <- read_csv("2015.csv", show_col_types = FALSE) |>
         clean_names()

```

## Exploratory Plot
```{r}
ggplot(happy, aes(economy_gdp_per_capita, happiness_score)) +
  geom_point() +
  geom_smooth() +
  labs(x = "GDP per Capita (scaled)",
       y = "Happiness Score")
```

## Data Tidying and Cleaning
```{r}
library(tidyverse)
library(janitor)   # clean_names()
library(skimr)     # quick data skim

happy <- read_csv("2015.csv", show_col_types = FALSE) |>
         clean_names() |>
         mutate(
           region = as.factor(region)          # treat Region as categorical
         ) |>
         drop_na()                             # 2015 data have no NAs, but good habit

skimr::skim(happy)
```

## Exploratory Data Analysis
```{r}
# 1) Summary stats by Region
eda_table <- happy |>
  group_by(region) |>
  summarise(
    countries       = n(),
    avg_happiness   = mean(happiness_score),
    avg_gdp         = mean(economy_gdp_per_capita),
    avg_life_expect = mean(health_life_expectancy),
    .groups = "drop"
  )

# 2) Boxplot of Happiness Score by Region
ggplot(happy, aes(region, happiness_score)) +
  geom_boxplot(fill = "grey80") +
  coord_flip() +
  labs(x = NULL, y = "Happiness Score",
       title = "How Happiness Varies by Region")

# 3) Correlation plot of numeric variables
library(GGally)
happy |>
  select(happiness_score,
         economy_gdp_per_capita,
         family,
         health_life_expectancy,
         freedom,
         trust_government_corruption,
         generosity) |>
  GGally::ggpairs(title = "Pair-wise Relationships Among Key Metrics")
```

## Appendix
```{r codeAppend, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}```
